---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# skynet

[![Build Status](https://travis-ci.org/FilipeamTeixeira/skynet.svg?branch=master)](https://travis-ci.org/FilipeamTeixeira/skynet.svg?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/skynet)](https://cran.r-project.org/package=skynet)
![](https://cranlogs.r-pkg.org/badges/grand-total/skynet?color=brightgreen)


# Overview

The rationale behind Skynet, is to provide researchers with a unifying tool overcoming some of the challenges faced when dealing with the Bureau of Transport Statistics, DB1B and T100 data.

## Installation

You can install skynet from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("FilipeamTeixeira/skynet")
```

## Example

To generate a directed network, you only need to type:

    library(skynet)
    import_db1b("folder/Coupon_2011Q1.csv", folder/Ticket_2011Q1.csv")
    make.netDir(OD_2011Q1, disp = TRUE, alpha = 0.05)

