---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# skynet

![Build Status](https://travis-ci.org/FilipeamTeixeira/skynet.svg?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/skynet)](https://cran.r-project.org/package=skynet)
![](https://cranlogs.r-pkg.org/badges/skynet?color=brightgreen)
[![Coverage status](https://codecov.io/gh/FilipeamTeixeira/Skynet/branch/master/graph/badge.svg)](https://codecov.io/github/FilipeamTeixeira/Skynet?branch=master)
[![](https://badges.ropensci.org/214_status.svg)](https://github.com/ropensci/onboarding/issues/214)

# Overview

The rationale behind Skynet, is to provide researchers with a unifying tool overcoming some of the challenges faced when dealing with the Bureau of Transport Statistics, DB1B and T100 data. 
The DB1B data consists of 2 sets of files, Coupon and Ticket. They can be both downloaded at https://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=289 and https://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=272 respectively while the T100 data can be found here https://www.transtats.bts.gov/Tables.asp?DB_ID=111.

## Note

To comply with R syntax guidelines, we changed to a clearer function naming from version 1.2.0. Deprecated functions are still present, but will be removed for the next versions.

## Installation

You can install skynet from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("FilipeamTeixeira/skynet")
```

## Import Data

To import data, simply type `import_db1b()` or `import_t100()` including the path to your desired file.  
**Note**: The Coupon file should take the first argument while the Ticket file should take the second argument.
    
```{r, eval=FALSE}
 library(skynet)
 import_db1b("folder/Coupon 2016Q1.csv", "folder/Ticket 2016Q1.csv")
 import_t100("folder/T100_2016.csv")
```

The BTS DB1B data consists of 2 sets of files, `Coupon` and `Ticket`.
They can be both downloaded at https://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=289 and https://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=272 respectively.

Despite being possible to download the complete zipped file, which includes all variables, due to its size, we recommend selecting the following set. 


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
knitr::kable(data.frame(Coupon = c("Itinerary ID", "Market ID", "Sequence Number", "Origin City Market ID",  
"Origin", "Year", "Quarter", "Destination City Market ID", "Destination", "Trip Break", "Operating Carrier", 
"Distance", "Gateway"),
Ticket = c("Itinerary ID", "Roundtrip", "Itinerary Yield", "Passengers",
"Itinerary Fare", "Bulkfare Indicator", "Distance","","","","","","")))
```

Since version 1.0.2 that the import method changed being the `netimport()` function no longer available.
When importing from the prezipped DB1B file, just add the argument `zip = TRUE` to the `import_db1b()` function. This does not apply to the T100 file which can be simply imported by typing `import_t100()`.
In order to save space, it is possible as well to import the prezipped file, and convert it to a smaller file with only the necessary variables, with the function `convert_raw()`. 


## Example

To generate a directed network, please type:

    library(skynet)
    # For DB1B data
    import_db1b("folder/Coupon_2011Q1.csv", "folder/Ticket_2011Q1.csv")
    make_net_dir(OD_2011Q1, disp = TRUE, alpha = 0.05)

    # For T100 data
    import_t100("folder/T100_2011.csv")
    make_net_dir(T100_2011Q1, disp = TRUE, alpha = 0.05)


